\pisodm: Building SOC Applications in the Presence of Non-Functional Requirements

### 2017-10-04 JournalPISODM2016/SubmittedV2/AnswersToReferees.tex

\begin{quotation}\sf\footnotesize

\noindent On the content:

1.	On Section 2, the related works are cited as such. This section lucks of an evaluation of these works, and of the motivation of your proposed method. 
\end{quotation}

\noindent 
We added 2 paragraphs at the end of Section 2: (i) considering the evaluation of the related work; and (ii) also motivating our proposed method. 
\begin{quotation}\sf\footnotesize

@@

2.	This method can be fully automated? Is it appropriate to non-IT professionals? 
\end{quotation}

\noindent 
We highlighted at the beginning of section 4, that the method is semi automatic, and we added a comment in as footnote explaining the reasons the method is not completely automatic.

@@

\begin{quotation}\sf\footnotesize

3.	While designing the application to develop using your models, how is the selection of services done? 
\end{quotation}

\noindent 
We highlighted in section 3.2.2, how the service is chosen.  To be chosen and executed the service must fit the contract (constraints). If the service fits the contract it will used in the execution flow.

@@

\noindent 
\hl{TO BE COMPLETED.}

\begin{quotation}\sf\footnotesize

6.	These non-functional requirements can include other types of constraints? 
\end{quotation}

\noindent 
As the non-functional requirements concept used in our method are related to business rules (constraints / contract rules) associated to the behavior of the application (services), it is possible include different types of constraints. The constraints are defined as pre- and post-conditions relates to services. So, during the design process, it is possible to include different kind of contracts related to services.  Figure 4 ($\pi$-UseCase Meta-Model.) presents the Policy View Concepts used to design the constraints.

@@

\begin{quotation}\sf\footnotesize

7.	On page 13, you specify that the model can be automatically translated into actual computer programs. Is this actually developed? If not, how is that can be possible? 
\end{quotation}

\noindent 
Yes, it is automatically translated into $\pi$-PEWS \cite{MPC08,BaCAM05,BHM06,BHM06rep} specification programs. It is possible because of the PIM to PSM transformation rules described in sections 4.3 and 4.4 (using Acceleo specification programs) and also described in \cite{placidoPhDThesis2012}.

### 2017-10-04 JournalPISODM2016/SubmittedV2/main.tex
Answer for the questions 1-3 and 6-7 to the Reviewer A.

Beside the existence of these works, it is important to remark that authors continue emphasizing the limited attention of the model-driven proposals for service development to NFRs~\cite{Ameller201542,Ameller2010}.%,SouzaNeto201684}. 
As Ameller et al. realize in a recently published literature review, there are \textit{``Not many papers that presented a generic approach to deal with NFRs"}. Those authors also remark the lack of industrial use cases~\cite{Ameller201542}. 
Our method to develop service-oriented applications aims to cover this lack of proposals dealing with NFRs and gives support to real needs, as presented in a company related to the risk assessment industry.

@@

 \begin{figure*}[h]
\center
\includegraphics[width=1\textwidth]{./figures/UseCaseMetaModel.pdf}
\caption{\label{fig:CIM:usecasemetamodel} $\pi$-UseCase Meta-Model.}
\end{figure*}

@@

This model shows the set of logically related activities to be performed in a service-based application.
The activities of this model represent a behavior that is part of the business logic of the application.
A \textit{$\pi$-ServiceProcess}  model contains three main elements: (i) service process, (ii) service activity and (iii) activity contract. A service activity represents an operation that is part of the execution flow, and it is modeled as an {\sc Action}.
An activity contract represents the {\sc Non-functional Requirement}  that is also part of the execution flow of a service, identified  as a stereotyped activity ({\sf assertion}).
The {\sc Assertion}s associated to an {\sc Action} compose a {\sc Contract}. 
Using the concepts {\sc Contract} and {\sc Assertion} it is possible to specify each activity service, by defining its pre- and post-conditions.

@@

\pisodm  proposes intra-level transformation rules from $\pi$-use case to $\pi$-service process models, $\pi$-service process to $\pi$-service composition  models, as well as inter-level transformation rules from $\pi$-service composition to $\pi$-PEWS models. 
These rules were implemented into a semi-automatic tool (the \pisodm  plugin for Eclipse).
The transformations from CIM to PIM level models are not automatized, due to the informal nature of the CIM level.

@@

\item 	We  used Acceleo\footnotemark \footnotetext{\hl{http://www.acceleo.org/pages/home/en}} to implement  the code generation plug-in to generate executable code. 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em
A-Policy} based service composition execution environment.
\end{compactitem}

>>>>

Beside the existence of these works, it is important to remark that authors continue emphasizing the limited attention of the model-driven proposals for service development to NFRs~\cite{Ameller201542,Ameller2010}.%,SouzaNeto201684}. 
As Ameller et al. realize in a recently published literature review, there are \textit{``Not many papers that presented a generic approach to deal with NFRs"}. Those authors also remark the lack of industrial use cases~\cite{Ameller201542}.

\hl{
To the best of our knowledge, and based on the related work analysis, there are no proposals that define a service-oriented approach for the whole development of systems, considering
non-functional requirements. Although the efforts that have been made to support the new
technological proposals for the Web such as web services, in general, these
development approaches focus their processes on traditional software
engineering, with emphasis on the functional aspects of the
application.} Our method to develop service-oriented applications aims to cover this lack of proposals dealing with NFRs and gives support to real needs, as presented in a company related to the risk assessment industry. \hl{Our approach targets to deal with characteristics of the 3
areas\footnote{(i) Non-Functional Requirements; (ii) Methodology for Service-Oriented Development;
and (iii) Model-Driven Engineering} in order to propose a method for
developing service-oriented applications considering non-functional aspects.

The motivation of our method is to target the specification of non-functional aspects in the early stages of development of service-oriented  applications. Our approach specifies \textit{policies} for modeling the non-functional requirements of a service separating the specification of non-functional requirements from the main functionalities of the application. This is interesting because once the policies for a given application have been defined they can be reused and/or specialized for another application with similar requirements.
}

@@

 \begin{figure*}[h]
\center
\includegraphics[width=1\textwidth]{./figures/UseCaseMetaModel.pdf}
\caption{\label{fig:CIM:usecasemetamodel} $\pi$-\hl{UseCase Meta-Model}.}
\end{figure*}

@@

\hl{This model shows the set of logically related activities to be performed in a service-based application.
The activities of this model represent a behavior that is part of the business logic of the application.}
A \textit{$\pi$-ServiceProcess}  model contains three main elements: (i) service process, (ii) service activity and (iii) activity contract. A service activity represents an operation that is part of the execution flow, and it is modeled as an {\sc Action}.
An activity contract represents the {\sc Non-functional Requirement}  that is also part of the execution flow of a service, identified  as a stereotyped activity ({\sf assertion}).
The {\sc Assertion}s associated to an {\sc Action} compose a {\sc Contract}. 
Using the concepts {\sc Contract} and {\sc Assertion} it is possible to specify each activity service, by defining its pre- and post-conditions. \hl{The service that fits the {\sc Contract} and its {\sc Assertions} will be chosen and used in the execution flow.}

@@

\pisodm  proposes intra-level transformation rules from $\pi$-use case to $\pi$-service process models, $\pi$-service process to $\pi$-service composition  models, as well as inter-level transformation rules from $\pi$-service composition to $\pi$-PEWS models. 
\hl{These rules were implemented into a semi-automatic tool} (the \pisodm  plugin\footnotemark \footnotetext{\hl{At each stage of transformation it is possible to include new constraints. In the more abstract stages of transformation there is complete automation in the models (i.e., from $\pi$-UseCase to $\pi$-ServiceProcess, and from $\pi$-ServiceProcess to $\pi$-ServiceComposition). On the other hand, to the transformation to specific platforms (i.e. from $\pi$-ServiceComposition to $\pi$-PEWS), a review by the analyst is required after the generation of the specification code. Thus, the method is more appropriate for those who have prior knowledge of Information Technology.}} for Eclipse).
The transformations from CIM to PIM level models are not automatized, due to the informal nature of the CIM level.

@@

\item 	\hl{We  used Acceleo}\footnotemark \footnotetext{\hl{http://www.acceleo.org/pages/home/en}} \hl{to implement  the code generation plug-in to generate executable code.} 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em
A-Policy} based service composition execution environment.
\end{compactitem}
